### 2020-04-14 paper/ICML/rebuttal.txt

Q2. The tools adopted is from classical tensor analysis (rank constrained M estimation) -This also can be related to reviewer #3’s question,

A2. I agree that rank constrained M estimation is common in previous works. In our paper, however, we established the theoretical results guaranteeing optimal convergence and complexity of M estimation. Those theories make our paper distinctive from others and justify using constrained M estimation.

Reviewer #4
Q1. The initialization of “b” is quite important to obtain good local optima. Do you have any good strategy to initialize b?(Initialization problem, may be the reviewer concerns identifiability issue too.)

A1.First, we perform multiple initialization methods to increase the probability of getting good initialization. Actual value of threshold parameter b cannot be decided by itself but with consideration of signal tensor and response variables, which means if signal tensor shifted left then b will follow (consider P(y\geq l) = f(b_l-theta)). Identifiability issue can arise from this point. So we set signal tensor sum to zero to solve the identifiability issue. Since we have mean 0 signal tensor, it is natiural for us to initialize the threshold parameter b is to set an increasing sequence such that the cumulative density of logisic distribution between consecutative threshold is the same. This initialization strategy for b combining with multiple initialization makes algorithm stable and start in good position to obtain good local optima.

### 2020-04-14 paper/ICML/rebuttal.txt

Q2. The tools adopted is from classical tensor analysis (rank constrained M estimation) -This also can be related to reviewer #3’s question,

A2. I agree that rank constrained M estimation is common in previous works. In our paper, however, we established the theoretical results guaranteeing optimal convergence and complexity of M estimation. Those theories make our paper distinctive from others and justify using constrained M estimation.

Reviewer #4
Q1. The initialization of “b” is quite important to obtain good local optima. Do you have any good strategy to initialize b?(Initialization problem, may be the reviewer concerns identifiability issue too.)

A1.First, we perform multiple initialization methods to increase the probability of getting good initialization. Actual value of threshold parameter b cannot be decided by itself but with consideration of signal tensor and response variables, which means if signal tensor shifted left then b will follow (consider P(y\geq l) = f(b_l-theta)). Identifiability issue can arise from this point. So we set signal tensor sum to zero to solve the identifiability issue. Since we have mean 0 signal tensor, it is natiural for us to initialize the threshold parameter b as an increasing sequence such that the cumulative density of the logistic distribution between consecutative threshold is the same (or an increasing seqeunece such that the cumulative density of the logistic distribution between consecutative threshold reflects the proportion of ordinal values in the data). This initialization strategy for b combining with multiple initialization makes algorithm stable and start in good position to obtain good local optima.

### 2020-04-14 paper/ICML/rebuttal.txt

Q2. Computational cost for the optimization.
MovieLens: dataset represents rating tuples of the form: <user, item, rating, timestamp>. There are diverse datasets such as 
100k:100,000 ratings from 1000 users on 1700 movies
1M:1 million ratings from 6000 users on 4000 movies
10M:10 million ratings and 100,000 tag applications applied to 10,000 movies by 72,000 users
20M:20 million ratings and 465,000 tag applications applied to 27,000 movies by 138,000 users
https://papers.nips.cc/paper/8860-singleshot-a-scalable-tucker-tensor-decomposition.pdf (Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS.)
https://arxiv.org/pdf/1710.02261v1.pdf(Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.)
: uses data where 6040 users, 100209 ratings from 1 to 6 on 3076 movies , between April,2000 and February,2003
(It assumes sparse tensor which means rank is really low)

A2. For bigger dataset, like Movielens dataset, we can use small rank to reduce computations like other papers have done for the application on Movielens such as
(Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS,
Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.) Here they used Tucker rank (5,5,5) and (10,10,10). This small rank makes implementation much faster (our brain dataset rank was (23,23,8) which is not quite small)
There are several rooms for us to improve our current algorithm for big data applications. Movielens datasets have more than 100000 ratings which makes computation of gradient challenging. However our loss function has good structure to apply stochastic gradient descent method. SGD is one way to improve our computation time. In addition, we can select a row-wise update method for factor matrices. In our alternating gradient based algorithm, we can show that all rows of factor matrices are independent of each other in the sense of minimizing our loss function. This property enables us to use parallel computing which significantly reduces memory requirement for updating factor matrices. I think those approaches enable us to apply big dataset like MovieLens.

### 2020-04-14 paper/ICML/rebuttal.txt

A2. For bigger dataset, like Movielens dataset, we can use small rank to reduce computations like other papers have done for the application on Movielens such as
(Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS,
Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.) Here they used Tucker rank (5,5,5) and (10,10,10). This small rank makes implementation much faster (our brain dataset rank was (23,23,8) which is not quite small)
There are several rooms for us to improve our current algorithm for big data applications. Movielens datasets have more than 100000 ratings which makes computation of gradient challenging. However our loss function has good structure to apply stochastic gradient descent method. SGD is one way to improve our computation time. In addition, we can select a row-wise update method for factor matrices. In our alternating gradient based algorithm, we can show that all rows of factor matrices are independent of each other in the sense of minimizing our loss function. This property enables us to use parallel computing which significantly reduces memory requirement for updating factor matrices. I think those approaches enable us to apply big dataset like MovieLens.

Reviewer #3
Q1. Can not find tractable algorithms for MLE ( He or she concerns about the gap between theoretical result and real application)

A1.
It is fair argument. However, no other approaches succeed to have tractable algorithm  to estimate global optimizer. The other methods include the nuclear norm approaches sacrificing rank constraints but still having computation hard algorithm. Also, our paper is meaningful in the sense that we show that this straightforward method works better than other approaches even with relaxed local minimum condition. Our paper shows theoretical reason for this outpeformance: MLE has optimal convergence rate and optimal complexity which makes even local minimum estimation outperforms the other methods with suboptimal rate. This paper provides one step closer to ultimate goal of tensor completion method for ordinal dataset.

In the matrix case of our optimization problem, many papers have shown that one would expect to achieve global convergence for the rank constrained optimization problem provided algorithms converge to a local minimum.
(R. Salakhutdinov and A. Mnih. Probabilistic matrix factorization. In Advances in Neural Information Processing Systems, 2008.,
T. Cai and W.-X. Zhou. A Max-Norm Constrained Minimization Approach to 1-Bit Matrix Completion. Journal of Machine Learning Research, 14:3619–3647, 2013.,
Ge, R., Jin, C., & Zheng, Y. (2017, August). No spurious local minima in nonconvex low rank problems: A unified geometric analysis. In Proceedings of the 34th International Conference on Machine Learning-Volume 70 (pp. 1233-1242). JMLR. org.)

In the line of those work, many studies try to show there is no spurious local minima in our tensor optimization problem.
(Haeffele, B. D., & Vidal, R. (2015). Global optimality in tensor factorization, deep learning, and beyond. arXiv preprint arXiv:1506.07540.,
Ge, R., & Ma, T. (2017). On the optimization landscape of tensor decompositions. In Advances in Neural Information Processing Systems (pp. 3653-3663).,
Sanjabi, M., Baharlouei, S., Razaviyayn, M., & Lee, J. D. (2019). When Does Non-Orthogonal Tensor Decomposition Have No Spurious Local Minima?. arXiv preprint arXiv:1911.09815.
). 
Even though the optimization problems in those papers are the special case of our problem (for example assuming that factor matrices are identical or adding extra regularization term), they are giving us positive conjecture that our algorithm may succeed finding global minima.

### 2020-04-14 paper/ICML/rebuttal.txt

A2. For bigger dataset, like Movielens dataset, we can use small rank to reduce computations like other papers have done for the application on Movielens such as
(Traore, A., Bérar, M., & Rakotomamonjy, A. (2019). Singleshot : a scalable Tucker tensor decomposition. NeurIPS,
Oh, S., Park, N., Sael, L., & Kang, U. (2018). Scalable Tucker Factorization for Sparse Tensors - Algorithms and Discoveries. 2018 IEEE 34th International Conference on Data Engineering (ICDE), 1120-1131.) Here they used Tucker rank (5,5,5) and (10,10,10). This small rank makes implementation much faster (our brain dataset rank was (23,23,8) which is not quite small)
There are several rooms for us to improve our current algorithm for big data applications. Movielens datasets have more than 100000 ratings which makes computation of gradient challenging. However our loss function has good structure to apply stochastic gradient descent method. SGD is one way to improve our computation time. In addition, we can select a row-wise update method for factor matrices. In our alternating gradient based algorithm, we can show that all rows of factor matrices are independent of each other in the sense of minimizing our loss function. This property enables us to use parallel computing which significantly reduces memory requirement for updating factor matrices. I think those approaches enable us to apply big dataset like MovieLens.

Reviewer #3
Q1. Can not find tractable algorithms for MLE ( He or she concerns about the gap between theoretical result and real application)

A1.
It is fair argument. However, I want to emphasize that our theorems are not only applicable to the global optimal MLE but any local optimums suffice as long as it has smaller loss function than at true parameter. This implies that the local optimums to the MLE have the optimal rate in the condition. Therefore, I would say that the optimality of the MLE is still useful. In the same line,  we argue that the finding a local minimum is suffice not only with experiments but also with theoretical analysis in our proof. 
The experiments are coincide with the theorems in the sense that our approach with possible local minimas outperforms other approaches (optimal vs suboptimal).

As you mentioned, the sufficiency of the finding local optima of nonconvex problems is very active area. Most of works are focusing on showing that local optima is global optima. Those approaches are different from our approach but wil provide another good point of our estimation method.

In the matrix case, many papers have shown that one would expect to achieve global convergence for the rank constrained optimization problem provided algorithms converge to a local minimum.
(R. Salakhutdinov and A. Mnih. Probabilistic matrix factorization. In Advances in Neural Information Processing Systems, 2008.,
T. Cai and W.-X. Zhou. A Max-Norm Constrained Minimization Approach to 1-Bit Matrix Completion. Journal of Machine Learning Research, 14:3619–3647, 2013.,
Ge, R., Jin, C., & Zheng, Y. (2017, August). No spurious local minima in nonconvex low rank problems: A unified geometric analysis. In Proceedings of the 34th International Conference on Machine Learning-Volume 70 (pp. 1233-1242). JMLR. org.)

In the tensor case, many studies try to show there is no spurious local minima for the rank constrained optimization problem.
(Haeffele, B. D., & Vidal, R. (2015). Global optimality in tensor factorization, deep learning, and beyond. arXiv preprint arXiv:1506.07540.,
Ge, R., & Ma, T. (2017). On the optimization landscape of tensor decompositions. In Advances in Neural Information Processing Systems (pp. 3653-3663).,
Sanjabi, M., Baharlouei, S., Razaviyayn, M., & Lee, J. D. (2019). When Does Non-Orthogonal Tensor Decomposition Have No Spurious Local Minima?. arXiv preprint arXiv:1911.09815.
). 
Even though the optimization problems in those papers are the special case of our problem (for example assuming that factor matrices are identical or adding extra regularization term), they are giving us positive conjecture that our algorithm may succeed finding global minima.

### 2020-04-19 paper/ICML/rebuttal0419.txt

We thank the reviewers for helpful comments and feedback. Our responses are detailed below.

To Reviewer 3:
* Regarding theoretical contributions and MLE approach

We highlight two main contributions that set our work apart from current literature. 
1) [New challenges with non-classical observations] The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known. The limited information makes the statistical recovery notably hard, not to mention the associated computation.

We are among the first to establish the recovery theory for both the signal tensors and the quantization operators from a limited number of highly discrete entries. Our model achieves invariance under a reversal of categories and consistency under merging/splitting of contiguous categories. A distinctive non-monotonic phase-transition pattern is demonstrated (Figure 2b). None of these features are ever considered in classical tensor analysis. Furthermore, we prove that, surprisingly, recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors. These results fill the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.

2) [New challenges with regularized MLE] The properties of regularized MLE are challenging in our tensor model. Current approaches for solving similar problems fall into two categories: convex (e.g., nuclear-norm regularized) MLE and non-convex MLE (our approach). Most work focuses on the former. However, in the tensor case, convex MLE suffers from both computational in-traceability and statistical sub-optimality. This is because computing nuclear norm is NP-hard for tensors.

We advocate the non-convex approach and provide strong evidence for its success in our setting. We improve the statistical error bound from d^{-(K-1)/2} to d^{-(K-1)}, which is substantial as the tensor order K increases. This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2019]. As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractably) convex MLE approach. While the computational limit remains open (as in many modern ML problems), we believe our improved statistical bound provides a useful benchmark for future algorithm development.

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convexity’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that the each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered. 
* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that applying to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations.'
* rank-constrained MLE
See our response to reviewer 3 regarding `New challenges with regularizeMLE.'

### 2020-04-19 paper/ICML/rebuttal0419.txt

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convexity’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that the each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered. 
* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that application to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations'
* rank-constrained MLE
See our response to reviewer 3 regarding `New challenges with regularizeMLE'

### 2020-04-19 paper/ICML/rebuttal0419.txt

2) [New challenges with regularized MLE] The properties of regularized MLE are challenging in our tensor model. Current approaches for solving similar problems fall into two categories: convex (e.g., nuclear-norm regularized) MLE and non-convex MLE (our approach). Most work focuses on the former. However, in the tensor case, convex MLE suffers from both computational in-traceability and statistical sub-optimality. This is because computing nuclear norm is NP-hard for tensors.

We advocate the non-convex approach and provide strong evidence for its success in our setting. We improve the statistical error bound from d^{-(K-1)/2} to d^{-(K-1)}, which is substantial as the tensor order K increases. This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2019]. As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractably) convex MLE approach. While the computational limit remains open (as in many modern ML problems), we believe our improved statistical bound provides a useful benchmark for future algorithm development.

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

@@

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that the each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered. 
* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that application to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations'
* rank-constrained MLE
See our response to reviewer 3 regarding `New challenges with regularized MLE'

### 2020-04-19 paper/ICML/rebuttal0419.txt

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that the each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered.

* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that application to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations'

* rank-constrained MLE
See our response to reviewer 3 regarding `New challenges with regularized MLE'

### 2020-04-19 paper/ICML/rebuttal0419.txt

To Reviewer 3:
* Regarding theoretical contributions and MLE approach

We highlight two main contributions that set our work apart from current literature. 
1) [New challenges with non-classical observations] The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known. The limited information makes the statistical recovery notably hard, not to mention the associated computation.

We are among the first to establish the recovery theory for both the signal tensors and the quantization operators from a limited number of highly discrete entries. Our model achieves invariance under a reversal of categories and consistency under merging/splitting of contiguous categories. A distinctive non-monotonic phase-transition pattern is demonstrated (Figure 2b). None of these features are ever considered in classical tensor analysis. Furthermore, we prove that, surprisingly, recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors. These results fill the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.

2) [New challenges with regularized MLE] The properties of regularized MLE are challenging in our tensor model. Current approaches for solving similar problems fall into two categories: convex (e.g., nuclear-norm regularized) MLE and non-convex MLE (our approach). Most work focuses on the former. However, in the tensor case, convex MLE suffers from both computational in-tractability and statistical sub-optimality. This is because computing nuclear norm is NP-hard for tensors.

@@

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm convergences to local optimums, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work open up new inquiry that allows more machine learning researchers to contribute to this field.

To Reviewer 4:
* Initialization of b
We initialize b such that each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered.

### 2020-04-19 paper/ICML/rebuttal0419.txt

We highlight two main contributions that set our work apart from current literature. 
1) [New challenges with non-classical observations] The problem of ordinal/quantized tensor is fundamentally more challenging than previously well-studied tensors for two reasons: (a) the entries do not belong to exponential family distribution, and (b) the observation contains much less information, because neither the underlying signal nor the quantization operator is known. The limited information makes the statistical recovery notably hard, not to mention the associated computation.

We are among the first to establish the recovery theory for both the signal tensors and the quantization operators from a limited number of highly discrete entries. Our model achieves invariance under a reversal of categories and consistency under merging/splitting of contiguous categories. A distinctive non-monotonic phase-transition pattern is demonstrated (Figure 2b). None of these features are ever considered in classical tensor analysis. Furthermore, we prove that, surprisingly, recovery from quantized tensors achieves equally good information-theoretical convergence as the continuous tensors. These results fill the gap between classical and non-classical (ordinal) observations, thereby greatly enriching the tensor-model literature.

2) [New challenges with regularized MLE] The properties of regularized MLE are challenging in our tensor model. Current approaches for solving similar problems fall into two categories: convex (e.g., nuclear-norm regularized) MLE and non-convex MLE (our approach). Most work focuses on the former. However, in the tensor case, convex MLE suffers from both computational in-tractability and statistical sub-optimality. This is because computing nuclear norm is NP-hard for tensors.

We advocate the non-convex approach and provide strong evidence for its success in our setting. We improve the statistical error bound from d^{-(K-1)/2} to d^{-(K-1)}, which is substantial as the tensor order K increases. This solves the earlier open question whether the square root can be removed in an unlimited computational setting [Ghadermarzy et al 2018]. As a by-effect, we demonstrate the statistical optimality of the non-convex approach over the (computationally intractable) convex MLE approach. While the computational limit remains open (as in many modern ML problems), we believe our improved statistical bound provides a useful benchmark for future algorithm development.

@@

* Regarding the optimization

We agree with the reviewer that MLE is computationally hard. While our main contributions lie in the aforementioned two aspects, we do provide useful results on optimization. We show that the alternating algorithm converges to stationary points, and we prove (in Thm 4.1 and Supplements) that the local optimums with reasonably high objective suffice for accurate recovery. In practice, we find that the estimation is robust under randomized initialization, which resembles the `benign non-convex’ phenomenon in literature. In special cases such as rank-1 or orthogonally decomposable tensors, we are able to leverage recent advances on optimization to obtain the asymptotical number of initializations for global guarantee, however, at the cost of more stringent assumptions on the underlying signal tensor. For this reason, we choose not to pursue this direction but leave framework general for the current work.

Tensor learning is a clear challenge for further research. We believe our work provides a (somewhat surprising) new idea for quantized tensor problem, and we obtain results that were previously difficult to establish. We hope the work opens up new inquiry that allows more machine learning researchers to contribute to this field.

—
To Reviewer 4:
* Initialization of b
We initialize b such that each of the estimated categories has roughly equal size. Furthermore, multiple initializations are utilized and we choose the one that maximizes the objective. We find that such initialization strategy gives stable results in the simulations and the two applications considered.

@@

* Computational costs
We choose these two datasets for fair comparison because they were used in the previous work. We agree that application to large datasets such as MovieLens would be interesting. In fact, our alternating approach lends itself well to this context. The sub-step for solving tensor factors could be paralleled and be boosted by fast algorithms such as stochastic gradient descent. We would be glad to include this application, space permitting.

—
To Reviewer 5:
* Difference to classical tensors
See our response to reviewer 3 regarding `New challenges with non-classical observations'
